---
title: "Cancer research explorer - China vs US"
author: "Sergio Valbuena"
date: "2025-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
```


```{r}
proj_dir <- file.path("C:", "Users", "svalb", "OneDrive", "Documentos", "GitHub", "cancer_studies_deployment", "public", "data")
```

```{r}
df <- read.csv(file.path(proj_dir, "articles_country_year_cancer_lite.csv"))

top10_cancers <- df %>% group_by(Cancer) %>% summarise(articles_cancer = sum(Articles))
top10_cancers <- (top10_cancers[order(top10_cancers$articles_cancer, decreasing = TRUE),] %>% filter(!Cancer == "Undetermined cancer"))[1:10,]

df_lite <- df %>% filter(Country %in% c("China", "United States"))
df_lite <- df_lite %>% filter(Cancer %in% top10_cancers$Cancer)




```{r}
ggplot(df_lite, aes(x=Year, y=Articles, colour = Country))+
  geom_point()
```



```{r}
df_stats <- df_lite %>% pivot_wider(names_from = Cancer, values_from = Articles)

df_stats$Mean <- rowMeans(df_stats[unique(df_lite$Cancer)])


for(row in 1:nrow(df_stats)){
  df_stats[row, "Q25"] <- quantile(as.numeric(df_stats[row, unique(df_lite$Cancer)]), probs = c(0.25))
  df_stats[row, "Q50"] <- quantile(as.numeric(df_stats[row, unique(df_lite$Cancer)]), probs = c(0.50))
  df_stats[row, "Q75"] <- quantile(as.numeric(df_stats[row, unique(df_lite$Cancer)]), probs = c(0.75))
}
```



```{r}
ggplot(df_stats, aes(x=Year, y=Q50))+
  geom_line(aes(colour = Country))+
  geom_ribbon(aes(ymin = Q25, ymax = Q75, fill = Country), alpha = .15) +
  xlim(1987, 2025)
```



```{r}
df_stats_2 <- df_lite %>% pivot_wider(names_from = Cancer, values_from = Articles)

df_stats_2$Mean <- rowMeans(df_stats_2[unique(df_lite$Cancer)])


for(row in 1:nrow(df_stats_2)){
  upper <- as.data.frame(t(df_stats[row,unique(df_lite$Cancer)])) %>% filter(V1 > as.numeric(df_stats[row, "Mean"]))
  lower <- as.data.frame(t(df_stats[row,unique(df_lite$Cancer)])) %>% filter(V1 < as.numeric(df_stats[row, "Mean"]))
  
  df_stats_2[row, "upper_SD"] <- sd(upper$V1)
  df_stats_2[row, "lower_SD"] <- sd(lower$V1)
}
```


```{r}
ggplot(df_stats_2, aes(x=Year, y=Mean))+
  geom_line(aes(colour = Country))+
  geom_ribbon(aes(ymin = Mean - lower_SD, ymax = Mean + upper_SD, fill = Country), alpha = .15) +
  xlim(1987, 2025)
```


```{r}
df_stats_3 <- df_lite %>% pivot_wider(names_from = Cancer, values_from = Articles)

df_stats_3$Mean <- rowMeans(df_stats_3[unique(df_lite$Cancer)])

for(row in 1:nrow(df_stats_3)){
  df_stats_3[row, "SD"] <- sd(as.numeric(df_stats_3[row, unique(df_lite$Cancer)]))
}
```



































